**The Heart can Lie: A Preliminary Investigation of the Role of Interoception and Theory of Mind in Deception**

### 2024-09-16 paper/manuscript.Rmd
R3

Two measures of ToM and its related constructs were administered. The Yoni Task [@shamay2007dissociable] is a behavioral task which assesses first and second-order ToM abilities in both cognitive and affective domains. Participants were presented with the face of a character named "Yoni", surrounded by four colored pictures of objects or faces - one in each corner of the screen. In total, each participant completed 101 trials - 49 trials assessing their affective ToM abilities, 37 trials assessing their cognitive ToM abilities and 15 control trials. During each trial, participants were instructed to respond based on specific corresponding cues such as the directions of Yoni's eye gaze, facial expressions etc., In the control trials, participants made judgements based on Yoni's physical context (physical ToM). More specifically, in first-order trials, participants were instructed to make inferences about Yoni's mental state with regards to the objects surrounding it (e.g., "Yoni is thinking of..." for cognitive ToM trials or "Yoni likes..." for affective ToM trials). In more complex second-order trials, participants had to correctly infer the interaction between Yoni and others' mental states (e.g., "Yoni is thinking of the fruit that ... wants" for cognitive ToM trials or "Yoni likes the fruit that ... likes" for affective ToM trials).

>>>>

Two measures of ToM and its related constructs were administered. The Yoni Task [@shamay2007dissociable] is a behavioral task which assesses first and second-order ToM abilities in both cognitive and affective domains. Participants were presented with the face of a character named "Yoni", surrounded by four colored pictures of objects or faces - one in each corner of the screen. In total, each participant completed 101 trials - 49 trials assessing their affective ToM abilities, 37 trials assessing their cognitive ToM abilities and 15 control trials (physical TOM). During each trial, participants were given an instruction (e.g., "Yoni is thinking of ..." or "Yoni loves ...") and a specific cue (e.g., the directions of Yoni's eye gaze or Yoni's facial expressions) which they used to choose the correct answer among the four options presented. Participants were instructed to respond as quickly as possible using the corresponding keys on the given keyboard. In the control trials (physical TOM), the instruction (e.g., "Yoni is close to ...") and the cue (e.g., physical distance between Yoni and the options) required participants to respond based on Yoni's physical context. Additionally, the instructions were changed to assess the first and second-order abilities for cognitive and affective TOM. In first-order TOM trials, participants were instructed to make inferences about Yoni's mental state with regards to the objects surrounding it (e.g., "Yoni is thinking of..." for cognitive ToM trials or "Yoni likes..." for affective ToM trials). In more complex second-order TOM trials, participants had to correctly infer the interaction between Yoni and others' mental states (e.g., "Yoni is thinking of the fruit that ... wants" for cognitive ToM trials or "Yoni likes the fruit that ... likes" for affective ToM trials).
